---
title: "phenotype_predictions"
output: html_document
date: "2023-03-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
setwd("/Users/eef52/Documents/Cornell/UAV-UGV/LSP_Analysis")
```


#Load libraries
```{r}
library(sommer)
library(psych)
library(ggplot2)
library(stringr)
library(tidyverse)
library(caTools)
require(nnet)
library(ggfortify)
library(dplyr)
library(Metrics)
library(BGLR)
library(caret)
library(Rfast)
```


#Read in data
```{r}
cv = 'CV1'
cv_version = ''
lsp_size = 512
equivdates = TRUE

df_lidar <- read.csv(paste0("LSP_files/CV_Experiments/", cv, "_labels_equiv/labels_z_lidar_NYH2_NYH3_2020-24_", cv, cv_version, "_", lsp_size, "_equiv.csv"))
df_msi <- read.csv(paste0("LSP_files/CV_Experiments/", cv, "_labels_equiv/labels_z_msi_NYH2_NYH3_2020-24_", cv, cv_version, "_", lsp_size, "_equiv.csv"))
df_VIs <- read.csv("VI_labels/labels_VI_NYH2_NYH3_2020-24.csv")
df_lidar <- df_lidar[,-c(1)]
df_msi <- df_msi[,-c(1)]
df_VIs <- df_VIs[,-c(1)]

#If CVBP or CV1, keep only data in validation set for autoencoder model
if(cv == 'CVBP' | cv == 'CV1'){
  df_lidar <- df_lidar[df_lidar$train_set == "val",]
  df_msi <- df_msi[df_msi$train_set == "val",]
}

#Remove test
df_lidar <- df_lidar[df_lidar$train_set != "test",]
df_msi <- df_msi[df_msi$train_set != "test",]

#Remove lidar 2020 and NYH3 2021
df_lidar <- df_lidar[!df_lidar$year == 2020,]
df_lidar <- df_lidar[!(df_lidar$year == 2021 & df_lidar$field == 'NYH3'),]

#Remove lidar senescence date
# df_lidar <- df_lidar[df_lidar$GDD_equiv != '2533-2626',]

#Constrain to same set of plots
df_lidar <- df_lidar[df_lidar$uniqueID %in% unique(df_msi$uniqueID), ]
df_msi <- df_msi[df_msi$uniqueID %in% unique(df_lidar$uniqueID), ]

#Add genotype_num variable (df_msi is more complete than df_lidar; ensure same genotype_num for each pedigree across datasets)
df_msi <- df_msi %>%
  mutate(genotype_num = as.numeric(factor(Pedigree))) %>%
  mutate(genotype_num = as.factor(genotype_num))
df_lidar <- df_lidar %>%
  left_join(unique(df_msi[,c("Pedigree", "genotype_num")]), by = c("Pedigree"))
```


#Remove outliers and discarded plots by environment
```{r}
#Get G2F information
labels2020 = read_csv("https://de.cyverse.org/anon-files/iplant/home/shared/commons_repo/curated/GenomesToFields_G2F_data_2020/a._2020_phenotypic_data/g2f_2020_phenotypic_clean_data.csv")
labels2021 = read_csv("https://de.cyverse.org/anon-files/iplant/home/shared/commons_repo/curated/GenomesToFields_G2F_data_2021/a._2021_phenotypic_data/g2f_2021_phenotypic_clean_data.csv")
labels2022 = read_csv("https://de.cyverse.org/anon-files/iplant/home/shared/commons_repo/curated/GenomesToFields_G2F_data_2022/a._2022_phenotypic_data/g2f_2022_phenotypic_clean_data.csv")
labels2023 = read.csv("get_labels/G2F_2023/G2F_2023_modified.csv", check.names = FALSE)
labels2024 = read.csv("get_labels/G2F_2024/G2F_2024_modified.csv", check.names = FALSE)

#Filter to NYH2 & NYH3
labels2020 <- labels2020[labels2020$`Field-Location` %in% c('NYH2', 'NYH3'), ]
labels2021 <- labels2021[labels2021$`Field-Location` %in% c('NYH2', 'NYH3'), ]
labels2022 <- labels2022[labels2022$`Field-Location` %in% c('NYH2', 'NYH3'), ]
labels2023 <- labels2023[labels2023$`Field-Location` %in% c('NYH2', 'NYH3'), ]
labels2024 <- labels2024[labels2024$`Field-Location` %in% c('NYH2', 'NYH3'), ]

#Create single df
labels2020$`Date Plot Planted [MM/DD/YY]` <- as.Date(labels2020$`Date Plot Planted [MM/DD/YY]`[1], '%m/%d/%Y')
labels2022$Source <- NA
labels2023 <- labels2023[,-c(1)]
labels2024 <- labels2024[,-c(1)]
labels_all <- rbind(labels2020, labels2021, labels2022, labels2023, labels2024)

#Rename columns
colnames(labels_all)[colnames(labels_all)=='Plot'] <- 'plot'
colnames(labels_all)[colnames(labels_all)=='Range'] <- 'range'
colnames(labels_all)[colnames(labels_all)=='Pass'] <- 'column'
colnames(labels_all)[colnames(labels_all)=='Anthesis [days]'] <- 'anthesis_days'
colnames(labels_all)[colnames(labels_all)=='Silking [days]'] <- 'silking_days'
colnames(labels_all)[colnames(labels_all)=='Plant Height [cm]'] <- 'plant_height'
colnames(labels_all)[colnames(labels_all)=='Ear Height [cm]'] <- 'ear_height'
colnames(labels_all)[colnames(labels_all)=='Stand Count [# of plants]'] <- 'stand_count'
colnames(labels_all)[colnames(labels_all)=='Root Lodging [# of plants]'] <- 'root_lodging'
colnames(labels_all)[colnames(labels_all)=='Stalk Lodging [# of plants]'] <- 'stalk_lodging'
colnames(labels_all)[colnames(labels_all)=='Grain Moisture [%]'] <- 'grain_moisture'
colnames(labels_all)[colnames(labels_all)=='Test Weight [lbs]'] <- 'test_weight'
colnames(labels_all)[colnames(labels_all)=='Plot Weight [lbs]'] <- 'plot_weight'
colnames(labels_all)[colnames(labels_all)=='Grain Yield (bu/A)'] <- 'grain_yield'
colnames(labels_all)[colnames(labels_all)=='Year'] <- 'year'
colnames(labels_all)[colnames(labels_all)=='Field-Location'] <- 'field'
labels_all <- as.data.frame(labels_all)
labels_all$plot <- as.integer(labels_all$plot)
labels_all$column <- as.integer(labels_all$column)

#Add ASI
labels_all$ASI <- labels_all$silking_days - labels_all$anthesis_days

#Add uniqueID
labels_all <- labels_all %>% unite("uniqueID", c("field", "year", "plot"), sep = "_", remove = FALSE, na.rm = FALSE)

phenos <- c("grain_yield", "anthesis_days", "silking_days", "plant_height", "ear_height", "stand_count", "root_lodging", "stalk_lodging", "grain_moisture", "plot_weight", "test_weight", "ASI")
for(pheno in phenos){
  print(pheno)
  for(year in 2020:2023){
    print(year)
    mean = mean(labels_all[labels_all$year == year, pheno], na.rm = TRUE)
    sd = sd(labels_all[labels_all$year == year, pheno], na.rm = TRUE)
    thresh_max = mean + (3*sd)
    thresh_min = mean - (3*sd)
    for(i in 1:nrow(labels_all)){
      if((labels_all[i, "year"] == year) & (is.na(labels_all[i, pheno]) == FALSE)){
        if(labels_all[i, pheno] > thresh_max || labels_all[i, pheno] < thresh_min){
          print(labels_all[i, 'plot'])
          labels_all[i, pheno] = NA
        }
        if(is.na(labels_all$`Plot Discarded [enter 'yes' or blank]`[i]) == FALSE){
          print(labels_all[i, 'plot'])
          labels_all[i, pheno] = NA
        }
      }
    }
  }
  #remove outliers by field for 2024
  for(year in c(2024)){
    print(year)
    for(field in c('NYH2', 'NYH3')){
      print(field)
      mean = mean(labels_all[labels_all$year == year & labels_all$field == field, pheno], na.rm = TRUE)
      sd = sd(labels_all[labels_all$year == year & labels_all$field == field, pheno], na.rm = TRUE)
      thresh_max = mean + (3*sd)
      thresh_min = mean - (3*sd)
      for(i in 1:nrow(labels_all)){
        if((labels_all[i, "year"] == year) & (labels_all[i, "field"] == field) & (is.na(labels_all[i, pheno]) == FALSE)){
          if(labels_all[i, pheno] > thresh_max || labels_all[i, pheno] < thresh_min){
            print(labels_all[i, 'plot'])
            labels_all[i, pheno] = NA
          }
          if(is.na(labels_all$`Plot Discarded [enter 'yes' or blank]`[i]) == FALSE){
            print(labels_all[i, 'plot'])
            labels_all[i, pheno] = NA
          }
        }
      }
    }
  }
}

#Reflect changes in df_lidar and df_msi
df_lidar <- df_lidar %>%
  left_join(labels_all %>% select(uniqueID, all_of(phenos)), by = "uniqueID", suffix = c("", ".df1"))
for(pheno in phenos){
  df_lidar[[pheno]] <- ifelse(is.na(df_lidar[[paste0(pheno, ".df1")]]), NA, df_lidar[[pheno]])
}
df_lidar <- df_lidar %>%
  select(-ends_with(".df1"))

df_msi <- df_msi %>%
  left_join(labels_all %>% select(uniqueID, all_of(phenos)), by = "uniqueID", suffix = c("", ".df1"))
for(pheno in phenos){
  df_msi[[pheno]] <- ifelse(is.na(df_msi[[paste0(pheno, ".df1")]]), NA, df_msi[[pheno]])
}
df_msi <- df_msi %>%
  select(-ends_with(".df1"))
```


#Create timeseries matrices
```{r}
#Get VIs for each GDD as new variable
VIs <- colnames(df_VIs %>% select(-"file_name"))
vi_timeseries <- df_msi %>%
  pivot_wider(
    names_from = GDD_equiv,   # The GDD values become column names
    values_from = VIs   # The VI values populate the new columns
  )
vi_timeseries <- cbind(vi_timeseries[,'uniqueID'], vi_timeseries[,setdiff(colnames(vi_timeseries), colnames(df_msi))])
vi_timeseries <- vi_timeseries %>% group_by_at(vars(uniqueID)) %>%
  summarise(across(where(is.numeric), ~mean(.x, na.rm = TRUE)))
vi_timeseries <- as.matrix(vi_timeseries)
rownames(vi_timeseries) <- vi_timeseries[,'uniqueID']
vi_timeseries <- vi_timeseries[,2:ncol(vi_timeseries)]
#Center and scale
storage.mode(vi_timeseries) <- "numeric"
vi_timeseries_scaled_by_lf <- scale(vi_timeseries, center = TRUE, scale = TRUE)

#Get MSI latent codes for each GDD as new variable
msi_timeseries <- df_msi %>%
  pivot_wider(
    names_from = GDD_equiv,   # The 'date' values become column names
    values_from = paste0("V", 1:lsp_size)   # The LSP values populate the new columns
  )
msi_timeseries <- cbind(msi_timeseries[,'uniqueID'], msi_timeseries[,setdiff(colnames(msi_timeseries), colnames(df_msi))])
msi_timeseries <- msi_timeseries %>% group_by_at(vars(uniqueID)) %>%
  summarise(across(where(is.numeric), ~mean(.x, na.rm = TRUE)))
msi_timeseries <- as.matrix(msi_timeseries)
rownames(msi_timeseries) <- msi_timeseries[,'uniqueID']
msi_timeseries <- msi_timeseries[,2:ncol(msi_timeseries)]
#Center and scale
storage.mode(msi_timeseries) <- "numeric"
msi_timeseries_scaled_by_lf <- scale(msi_timeseries, center = TRUE, scale = TRUE)

#Get lidar latent codes for each GDD as new variable
lidar_timeseries <- df_lidar %>%
  pivot_wider(
    names_from = GDD_equiv,   # The 'date' values become column names
    values_from = paste0("V", 1:lsp_size)   # The LSP values populate the new columns
  )
lidar_timeseries <- cbind(lidar_timeseries[,'uniqueID'], lidar_timeseries[,setdiff(colnames(lidar_timeseries), colnames(df_lidar))])
lidar_timeseries <- lidar_timeseries %>% group_by_at(vars(uniqueID)) %>%
  summarise(across(where(is.numeric), ~mean(.x, na.rm = TRUE)))
lidar_timeseries <- as.matrix(lidar_timeseries)
rownames(lidar_timeseries) <- lidar_timeseries[,'uniqueID']
lidar_timeseries <- lidar_timeseries[,2:ncol(lidar_timeseries)]
# impute missing data by GDD_equiv & year
year_from_row <- function(x) sub("^[^_]+_([0-9]{4})_.*$", "\\1", x)  # keep year
rn <- rownames(lidar_timeseries) # get rownames
year <- vapply(rn, year_from_row, character(1)) # get row indices of each year
groups_year <- split(seq_along(rn), year) # group row indices by year label
# for each year and each GDD_equiv, replace NA with the mean of that group
for(i in colnames(lidar_timeseries)){
  for(g in groups_year){
    vals <- as.numeric(lidar_timeseries[g, i])
    m <- mean(vals, na.rm = TRUE)
    if(is.finite(m)){
      nas <- is.na(vals)
      if(any(nas)) lidar_timeseries[g, i][nas] <- m
    }
  }
}
#Center and scale
storage.mode(lidar_timeseries) <- "numeric"
lidar_timeseries_scaled_by_lf <- scale(lidar_timeseries, center = TRUE, scale = TRUE)
lidar_timeseries_scaled_by_lf <- lidar_timeseries_scaled_by_lf[, colSums(is.na(lidar_timeseries_scaled_by_lf)) == 0]

#Get integrated latent codes for all dates
integrated_timeseries <- merge(msi_timeseries, lidar_timeseries, 
                          by = 'row.names', all = FALSE)
rownames(integrated_timeseries) <- integrated_timeseries[,'Row.names']
integrated_timeseries <- as.matrix(integrated_timeseries[,2:ncol(integrated_timeseries)])
#Center and scale
storage.mode(integrated_timeseries) <- "numeric"
integrated_timeseries_scaled_by_lf <- scale(integrated_timeseries, center = TRUE, scale = TRUE)
```


#Early season predictions
```{r}
#Early season predictions
#remove final lidar window
integrated_timeseries_scaled_by_lf <- integrated_timeseries_scaled_by_lf[,!grepl("2533-2626", colnames(integrated_timeseries_scaled_by_lf))]
lidar_timeseries_scaled_by_lf <- lidar_timeseries_scaled_by_lf[,!grepl("2533-2626", colnames(lidar_timeseries_scaled_by_lf))]
#remove next msi window
integrated_timeseries_scaled_by_lf <- integrated_timeseries_scaled_by_lf[,!grepl("2288-2477", colnames(integrated_timeseries_scaled_by_lf))]
msi_timeseries_scaled_by_lf <- msi_timeseries_scaled_by_lf[,!grepl("2288-2477", colnames(msi_timeseries_scaled_by_lf))]
vi_timeseries_scaled_by_lf <- vi_timeseries_scaled_by_lf[,!grepl("2288-2477", colnames(vi_timeseries_scaled_by_lf))]
#remove next msi window
integrated_timeseries_scaled_by_lf <- integrated_timeseries_scaled_by_lf[,!grepl("1943-2035", colnames(integrated_timeseries_scaled_by_lf))]
msi_timeseries_scaled_by_lf <- msi_timeseries_scaled_by_lf[,!grepl("1943-2035", colnames(msi_timeseries_scaled_by_lf))]
vi_timeseries_scaled_by_lf <- vi_timeseries_scaled_by_lf[,!grepl("1943-2035", colnames(vi_timeseries_scaled_by_lf))]
#remove next lidar window
integrated_timeseries_scaled_by_lf <- integrated_timeseries_scaled_by_lf[,!grepl("1267-1405", colnames(integrated_timeseries_scaled_by_lf))]
lidar_timeseries_scaled_by_lf <- lidar_timeseries_scaled_by_lf[,!grepl("1267-1405", colnames(lidar_timeseries_scaled_by_lf))]
#remove next msi window
integrated_timeseries_scaled_by_lf <- integrated_timeseries_scaled_by_lf[,!grepl("1008-1174", colnames(integrated_timeseries_scaled_by_lf))]
msi_timeseries_scaled_by_lf <- msi_timeseries_scaled_by_lf[,!grepl("1008-1174", colnames(msi_timeseries_scaled_by_lf))]
vi_timeseries_scaled_by_lf <- vi_timeseries_scaled_by_lf[,!grepl("1008-1174", colnames(vi_timeseries_scaled_by_lf))]
```


#Model Comparison - Predict Pheno from LSPs using Bayesian Regression - Season-Level CVBP
```{r}
run_BGLR <- function(cov_matrix, phenoDat, saveAt){
  #create folds for 5-fold cross-validation
  folds <- createFolds(phenoDat$pheno, k = 5)
  #set up a vector to store predicted values
  Ypred = rep(0, nrow(phenoDat))
  #loop for 5 fold cross validation
  for(i in 1:5){
    #setting up a vector of phenotypes for each fold
    Y = as.matrix(phenoDat$pheno)
    #mask data for prediction
    Y[folds[[i]]] <- NA
    #set up ETA list
    #Model <- list(list(K = cov_matrix, model = "RKHS"))
    Model <- list(list(X = cov_matrix, model = "BRR"))
    result <- BGLR(Y, response_type = "gaussian", ETA = Model, nIter = 20000, burnIn = 500, thin = 100, verbose = FALSE, saveAt = paste0(saveAt, "fold", i, "_"))
    #get pheno estimates from result list
    Yhat = result$yHat
    Ypred[folds[[i]]] = Yhat[folds[[i]]]
  }
  #get pheno
  Y = as.matrix(phenoDat$pheno)
  #get df of pred and actual phenotypes
  df = data.frame(pred = Ypred, actual = Y, uniqueID = as.matrix(phenoDat$uniqueID))
  #correlation between pred and actual
  cor = cor(Ypred, Y, use = 'complete.obs')
  print(cor)
  #rmse between pred and actual
  rmse = sqrt(mean((Y - Ypred)^2))
  print(rmse)
  #return correlation and rmse
  return(c(cor, rmse))
  #return(df)
}

#Prediction from VIs
phenos_list <- c("silking_days", "anthesis_days", "grain_yield", "plant_height", "ear_height", "stand_count", "stalk_lodging", "grain_moisture", "ASI")
pred_results_VI <- c()
for(pheno in phenos_list){
  set.seed(10)
  print(pheno)
  #get phenotype
  phenoDat <- df_msi[,c("uniqueID", pheno)]
  colnames(phenoDat) <- c("uniqueID", "pheno")
  #remove NAs from phenoDat
  phenoDat <- phenoDat[!is.na(phenoDat$pheno),]
  #get phenotype for each uniqueID
  phenoDat <- phenoDat %>% group_by_at(vars(uniqueID)) %>%
    summarize_each(funs(mean(., na.rm = TRUE)))
  #remove uniqueIDs where pheno is NA from cov mat
  cov_mat <- vi_timeseries_scaled_by_lf[(rownames(vi_timeseries_scaled_by_lf) %in% phenoDat$uniqueID), ]
  #sort pheno by uniqueID of cov mat & remove uniqueIDs not in cov mat
  rownames(phenoDat) <- phenoDat$uniqueID
  phenoDat <- phenoDat[rownames(cov_mat), ]
  #mean center phenotype by year
  sub_2020 <- grepl("2020", phenoDat$uniqueID)
  sub_2021 <- grepl("2021", phenoDat$uniqueID)
  sub_2022 <- grepl("2022", phenoDat$uniqueID)
  sub_2023 <- grepl("2023", phenoDat$uniqueID)
  mean_val_2020 <- mean(phenoDat$pheno[sub_2020], na.rm = TRUE)
  mean_val_2021 <- mean(phenoDat$pheno[sub_2021], na.rm = TRUE)
  mean_val_2022 <- mean(phenoDat$pheno[sub_2022], na.rm = TRUE)
  mean_val_2023 <- mean(phenoDat$pheno[sub_2023], na.rm = TRUE)
  phenoDat$pheno[sub_2020] <- phenoDat$pheno[sub_2020] - mean_val_2020
  phenoDat$pheno[sub_2021] <- phenoDat$pheno[sub_2021] - mean_val_2021
  phenoDat$pheno[sub_2022] <- phenoDat$pheno[sub_2022] - mean_val_2022
  phenoDat$pheno[sub_2023] <- phenoDat$pheno[sub_2023] - mean_val_2023
  for(i in 1:10){
    save_at <- paste0("BGLR_output/CVBP_dtype_comparisons_df/VI_", pheno, "_", cv, "_", i, "_")
    bglr_mod <- run_BGLR(cov_mat, phenoDat, save_at)
    pred_results <- data.frame(CV = cv, LatentDimSize = NA, EquivDates = equivdates, Pheno = pheno, Predictor = "Vegetation Indices", Correlation = bglr_mod[1], RMSE = bglr_mod[2])
    # pred_results <- bglr_mod
    # pred_results$pheno <- pheno
    pred_results_VI <- rbind(pred_results_VI, pred_results)
  }
}

write.csv(pred_results_VI, paste0("pred_results/CV_dtype_comparisons_df/pred_VI_", cv, ".csv"))

#Prediction from MSI latent codes
phenos_list <- c("silking_days", "anthesis_days", "grain_yield", "plant_height", "ear_height", "stand_count", "stalk_lodging", "grain_moisture", "ASI")
pred_results_msi <- c()
for(pheno in phenos_list){
  set.seed(10)
  print(pheno)
  #get phenotype
  phenoDat <- df_msi[,c("uniqueID", pheno)]
  colnames(phenoDat) <- c("uniqueID", "pheno")
  #remove NAs from phenoDat
  phenoDat <- phenoDat[!is.na(phenoDat$pheno),]
  #get phenotype for each uniqueID
  phenoDat <- phenoDat %>% group_by_at(vars(uniqueID)) %>%
    summarize_each(funs(mean(., na.rm = TRUE)))
  #remove uniqueIDs where pheno is NA from cov mat
  cov_mat <- msi_timeseries_scaled_by_lf[(rownames(msi_timeseries_scaled_by_lf) %in% phenoDat$uniqueID), ]
  #sort pheno by uniqueID of cov mat & remove uniqueIDs not in cov mat
  rownames(phenoDat) <- phenoDat$uniqueID
  phenoDat <- phenoDat[rownames(cov_mat), ]
  #mean center phenotype by year
  sub_2020 <- grepl("2020", phenoDat$uniqueID)
  sub_2021 <- grepl("2021", phenoDat$uniqueID)
  sub_2022 <- grepl("2022", phenoDat$uniqueID)
  sub_2023 <- grepl("2023", phenoDat$uniqueID)
  mean_val_2020 <- mean(phenoDat$pheno[sub_2020], na.rm = TRUE)
  mean_val_2021 <- mean(phenoDat$pheno[sub_2021], na.rm = TRUE)
  mean_val_2022 <- mean(phenoDat$pheno[sub_2022], na.rm = TRUE)
  mean_val_2023 <- mean(phenoDat$pheno[sub_2023], na.rm = TRUE)
  phenoDat$pheno[sub_2020] <- phenoDat$pheno[sub_2020] - mean_val_2020
  phenoDat$pheno[sub_2021] <- phenoDat$pheno[sub_2021] - mean_val_2021
  phenoDat$pheno[sub_2022] <- phenoDat$pheno[sub_2022] - mean_val_2022
  phenoDat$pheno[sub_2023] <- phenoDat$pheno[sub_2023] - mean_val_2023
  for(i in 1:10){
    save_at <- paste0("BGLR_output/CVBP_dtype_comparisons_df/msi_", pheno, "_", cv, "_", lsp_size, "_", i, "_")
    bglr_mod <- run_BGLR(cov_mat, phenoDat, save_at)
    pred_results <- data.frame(CV = cv, LatentDimSize = lsp_size, EquivDates = equivdates, Pheno = pheno, Predictor = "MSI Latent Phenotypes", Correlation = bglr_mod[1], RMSE = bglr_mod[2])
    # pred_results <- bglr_mod
    # pred_results$pheno <- pheno
    pred_results_msi <- rbind(pred_results_msi, pred_results)
  }
}

write.csv(pred_results_msi, paste0("pred_results/CV_dtype_comparisons_df/pred_msi_", cv, "_", lsp_size, ".csv"))

#Prediction from lidar latent codes
phenos_list <- c("silking_days", "anthesis_days", "grain_yield", "plant_height", "ear_height", "stand_count", "stalk_lodging", "grain_moisture", "ASI")
pred_results_lidar <- c()
for(pheno in phenos_list){
  set.seed(10)
  print(pheno)
  #get phenotype
  phenoDat <- df_lidar[,c("uniqueID", pheno)]
  colnames(phenoDat) <- c("uniqueID", "pheno")
  #remove NAs from phenoDat
  phenoDat <- phenoDat[!is.na(phenoDat$pheno),]
  #get phenotype for each uniqueID
  phenoDat <- phenoDat %>% group_by_at(vars(uniqueID)) %>%
    summarize_each(funs(mean(., na.rm = TRUE)))
  #remove uniqueIDs where pheno is NA from cov mat
  cov_mat <- lidar_timeseries_scaled_by_lf[(rownames(lidar_timeseries_scaled_by_lf) %in% phenoDat$uniqueID), ]
  #sort pheno by uniqueID of cov mat & remove uniqueIDs not in cov mat
  rownames(phenoDat) <- phenoDat$uniqueID
  phenoDat <- phenoDat[rownames(cov_mat), ]
  #mean center phenotype by year
  sub_2020 <- grepl("2020", phenoDat$uniqueID)
  sub_2021 <- grepl("2021", phenoDat$uniqueID)
  sub_2022 <- grepl("2022", phenoDat$uniqueID)
  sub_2023 <- grepl("2023", phenoDat$uniqueID)
  mean_val_2020 <- mean(phenoDat$pheno[sub_2020], na.rm = TRUE)
  mean_val_2021 <- mean(phenoDat$pheno[sub_2021], na.rm = TRUE)
  mean_val_2022 <- mean(phenoDat$pheno[sub_2022], na.rm = TRUE)
  mean_val_2023 <- mean(phenoDat$pheno[sub_2023], na.rm = TRUE)
  phenoDat$pheno[sub_2020] <- phenoDat$pheno[sub_2020] - mean_val_2020
  phenoDat$pheno[sub_2021] <- phenoDat$pheno[sub_2021] - mean_val_2021
  phenoDat$pheno[sub_2022] <- phenoDat$pheno[sub_2022] - mean_val_2022
  phenoDat$pheno[sub_2023] <- phenoDat$pheno[sub_2023] - mean_val_2023
  for(i in 1:10){
    save_at <- paste0("BGLR_output/CVBP_dtype_comparisons_df/lidar_", pheno, "_", cv, "_", lsp_size, "_", i, "_")
    bglr_mod <- run_BGLR(cov_mat, phenoDat, save_at)
    pred_results <- data.frame(CV = cv, LatentDimSize = lsp_size, EquivDates = equivdates, Pheno = pheno, Predictor = "LiDAR Latent Phenotypes", Correlation = bglr_mod[1], RMSE = bglr_mod[2])
    # pred_results <- bglr_mod
    # pred_results$pheno <- pheno
    pred_results_lidar <- rbind(pred_results_lidar, pred_results)
  }
}

write.csv(pred_results_lidar, paste0("pred_results/CV_dtype_comparisons_df/pred_lidar_", cv, "_", lsp_size, ".csv"))

#Prediction from integrated latent codes
phenos_list <- c("silking_days", "anthesis_days", "grain_yield", "plant_height", "ear_height", "stand_count", "stalk_lodging", "grain_moisture", "ASI")
pred_results_int <- c()
for(pheno in phenos_list){
  set.seed(10)
  print(pheno)
  #get phenotype
  phenoDat <- df_msi[,c("uniqueID", pheno)]
  colnames(phenoDat) <- c("uniqueID", "pheno")
  #remove NAs from phenoDat
  phenoDat <- phenoDat[!is.na(phenoDat$pheno),]
  #get phenotype for each uniqueID
  phenoDat <- phenoDat %>% group_by_at(vars(uniqueID)) %>%
    summarize_each(funs(mean(., na.rm = TRUE)))
  #remove uniqueIDs where pheno is NA from cov mat
  cov_mat <- integrated_timeseries_scaled_by_lf[(rownames(integrated_timeseries_scaled_by_lf) %in% phenoDat$uniqueID), ]
  #sort pheno by uniqueID of cov mat & remove uniqueIDs not in cov mat
  rownames(phenoDat) <- phenoDat$uniqueID
  phenoDat <- phenoDat[rownames(cov_mat), ]
  #mean center phenotype by year
  sub_2020 <- grepl("2020", phenoDat$uniqueID)
  sub_2021 <- grepl("2021", phenoDat$uniqueID)
  sub_2022 <- grepl("2022", phenoDat$uniqueID)
  sub_2023 <- grepl("2023", phenoDat$uniqueID)
  mean_val_2020 <- mean(phenoDat$pheno[sub_2020], na.rm = TRUE)
  mean_val_2021 <- mean(phenoDat$pheno[sub_2021], na.rm = TRUE)
  mean_val_2022 <- mean(phenoDat$pheno[sub_2022], na.rm = TRUE)
  mean_val_2023 <- mean(phenoDat$pheno[sub_2023], na.rm = TRUE)
  phenoDat$pheno[sub_2020] <- phenoDat$pheno[sub_2020] - mean_val_2020
  phenoDat$pheno[sub_2021] <- phenoDat$pheno[sub_2021] - mean_val_2021
  phenoDat$pheno[sub_2022] <- phenoDat$pheno[sub_2022] - mean_val_2022
  phenoDat$pheno[sub_2023] <- phenoDat$pheno[sub_2023] - mean_val_2023
  for(i in 1:10){
    save_at <- paste0("BGLR_output/CVBP_dtype_comparisons_df/int_", pheno, "_", cv, "_", lsp_size, "_", i, "_")
    bglr_mod <- run_BGLR(cov_mat, phenoDat, save_at)
    pred_results <- data.frame(CV = cv, LatentDimSize = lsp_size, EquivDates = equivdates, Pheno = pheno, Predictor = "Integrated Latent Phenotypes", Correlation = bglr_mod[1], RMSE = bglr_mod[2])
    # pred_results <- bglr_mod
    # pred_results$pheno <- pheno
    pred_results_int <- rbind(pred_results_int, pred_results)
  }
}

write.csv(pred_results_int, paste0("pred_results/CV_dtype_comparisons_df/pred_int_", cv, "_", lsp_size, ".csv"))

pred_results_all <- rbind(pred_results_msi, pred_results_lidar, pred_results_int)

write.csv(pred_results_all, "pred_results/pred_all_", cv, "_", lsp_size, ".csv")
```


#Model Comparison - Predict Pheno from LSPs using Bayesian Regression - Season-Level CV1
```{r}
run_BGLR <- function(cov_matrix, phenoDat, saveAt){
  #create 5 folds on genotypes (not rows)
  geno_folds <- createFolds(phenoDat$genotype_num, k = 5)
  #vector to store predicted values for each row
  Ypred <- rep(0, nrow(phenoDat))
  #loop for 5 fold cross validation
  for(i in 1:5){
    #full phenotype vector
    Y <- as.matrix(phenoDat$pheno)
    #mask test rows for prediction
    Y[geno_folds[[i]]] <- NA
    #set up ETA list
    Model <- list(list(X = cov_matrix, model = "BRR"))
    result <- BGLR(Y, response_type = "gaussian", ETA = Model, nIter = 20000, burnIn = 500, thin = 100, verbose = FALSE, saveAt = paste0(saveAt, "fold", i, "_"))
    #get pheno estimates from result list
    Yhat = result$yHat
    Ypred[geno_folds[[i]]] = Yhat[geno_folds[[i]]]
  }
  #get pheno
  Y = as.matrix(phenoDat$pheno)
  #get df of pred and actual phenotypes
  df = data.frame(pred = Ypred, actual = Y, uniqueID = as.matrix(phenoDat$uniqueID))
  #correlation between pred and actual
  cor = cor(Ypred, Y, use = 'complete.obs')
  print(cor)
  #rmse between pred and actual
  rmse = sqrt(mean((Y - Ypred)^2))
  print(rmse)
  #return correlation and rmse
  return(c(cor, rmse))
  #return(df)
}

#Prediction from VIs
phenos_list <- c("silking_days", "anthesis_days", "grain_yield", "plant_height", "ear_height", "stand_count", "stalk_lodging", "grain_moisture", "ASI")
pred_results_VI <- c()
for(pheno in phenos_list){
  set.seed(10)
  print(pheno)
  #get phenotype
  phenoDat <- df_msi[,c("uniqueID", pheno, "genotype_num")]
  colnames(phenoDat) <- c("uniqueID", "pheno", "genotype_num")
  #remove NAs from phenoDat
  phenoDat <- phenoDat[!is.na(phenoDat$pheno),]
  #get phenotype for each uniqueID
  phenoDat <- phenoDat %>% group_by_at(vars(uniqueID)) %>%
    summarise(across(where(is.numeric), ~mean(.x, na.rm = TRUE)), across(!where(is.numeric), ~first(.x)))
  #remove uniqueIDs where pheno is NA from cov mat
  cov_mat <- vi_timeseries_scaled_by_lf[(rownames(vi_timeseries_scaled_by_lf) %in% phenoDat$uniqueID), ]
  #sort pheno by uniqueID of cov mat & remove uniqueIDs not in cov mat
  rownames(phenoDat) <- phenoDat$uniqueID
  phenoDat <- phenoDat[rownames(cov_mat), ]
  #mean center phenotype by year
  sub_2020 <- grepl("2020", phenoDat$uniqueID)
  sub_2021 <- grepl("2021", phenoDat$uniqueID)
  sub_2022 <- grepl("2022", phenoDat$uniqueID)
  sub_2023 <- grepl("2023", phenoDat$uniqueID)
  mean_val_2020 <- mean(phenoDat$pheno[sub_2020], na.rm = TRUE)
  mean_val_2021 <- mean(phenoDat$pheno[sub_2021], na.rm = TRUE)
  mean_val_2022 <- mean(phenoDat$pheno[sub_2022], na.rm = TRUE)
  mean_val_2023 <- mean(phenoDat$pheno[sub_2023], na.rm = TRUE)
  phenoDat$pheno[sub_2020] <- phenoDat$pheno[sub_2020] - mean_val_2020
  phenoDat$pheno[sub_2021] <- phenoDat$pheno[sub_2021] - mean_val_2021
  phenoDat$pheno[sub_2022] <- phenoDat$pheno[sub_2022] - mean_val_2022
  phenoDat$pheno[sub_2023] <- phenoDat$pheno[sub_2023] - mean_val_2023
  for(i in 1:10){
    save_at <- paste0("BGLR_output/CV1_dtype_comparisons/VI_", pheno, "_", cv, "_", i, "_")
    bglr_mod <- run_BGLR(cov_mat, phenoDat, save_at)
    pred_results <- data.frame(CV = cv, LatentDimSize = NA, EquivDates = equivdates, Pheno = pheno, Predictor = "Vegetation Indices", Correlation = bglr_mod[1], RMSE = bglr_mod[2])
    # pred_results <- bglr_mod
    # pred_results$pheno <- pheno
    pred_results_VI <- rbind(pred_results_VI, pred_results)
  }
}

write.csv(pred_results_VI, paste0("pred_results/CV_dtype_comparisons/pred_VI_", cv, ".csv"))

#Prediction from MSI latent codes
phenos_list <- c("silking_days", "anthesis_days", "grain_yield", "plant_height", "ear_height", "stand_count", "stalk_lodging", "grain_moisture", "ASI")
pred_results_msi <- c()
for(pheno in phenos_list){
  set.seed(10)
  print(pheno)
  #get phenotype
  phenoDat <- df_msi[,c("uniqueID", pheno, "genotype_num")]
  colnames(phenoDat) <- c("uniqueID", "pheno", "genotype_num")
  #remove NAs from phenoDat
  phenoDat <- phenoDat[!is.na(phenoDat$pheno),]
  #get phenotype for each uniqueID
  phenoDat <- phenoDat %>% group_by_at(vars(uniqueID)) %>%
    summarise(across(where(is.numeric), ~mean(.x, na.rm = TRUE)), across(!where(is.numeric), ~first(.x)))
  #remove uniqueIDs where pheno is NA from cov mat
  cov_mat <- msi_timeseries_scaled_by_lf[(rownames(msi_timeseries_scaled_by_lf) %in% phenoDat$uniqueID), ]
  #sort pheno by uniqueID of cov mat & remove uniqueIDs not in cov mat
  rownames(phenoDat) <- phenoDat$uniqueID
  phenoDat <- phenoDat[rownames(cov_mat), ]
  #mean center phenotype by year
  sub_2020 <- grepl("2020", phenoDat$uniqueID)
  sub_2021 <- grepl("2021", phenoDat$uniqueID)
  sub_2022 <- grepl("2022", phenoDat$uniqueID)
  sub_2023 <- grepl("2023", phenoDat$uniqueID)
  mean_val_2020 <- mean(phenoDat$pheno[sub_2020], na.rm = TRUE)
  mean_val_2021 <- mean(phenoDat$pheno[sub_2021], na.rm = TRUE)
  mean_val_2022 <- mean(phenoDat$pheno[sub_2022], na.rm = TRUE)
  mean_val_2023 <- mean(phenoDat$pheno[sub_2023], na.rm = TRUE)
  phenoDat$pheno[sub_2020] <- phenoDat$pheno[sub_2020] - mean_val_2020
  phenoDat$pheno[sub_2021] <- phenoDat$pheno[sub_2021] - mean_val_2021
  phenoDat$pheno[sub_2022] <- phenoDat$pheno[sub_2022] - mean_val_2022
  phenoDat$pheno[sub_2023] <- phenoDat$pheno[sub_2023] - mean_val_2023
  for(i in 1:10){
    save_at <- paste0("BGLR_output/CV1_cv_comparisons/msi_", pheno, "_", cv, "_", lsp_size, "_", i, "_")
    bglr_mod <- run_BGLR(cov_mat, phenoDat, save_at)
    pred_results <- data.frame(CV = cv, LatentDimSize = lsp_size, EquivDates = equivdates, Pheno = pheno, Predictor = "MSI Latent Phenotypes", Correlation = bglr_mod[1], RMSE = bglr_mod[2])
    # pred_results <- bglr_mod
    # pred_results$pheno <- pheno
    pred_results_msi <- rbind(pred_results_msi, pred_results)
  }
}

write.csv(pred_results_msi, paste0("pred_results/CV_comparisons/pred_msi_", cv, "_", lsp_size, ".csv"))

#Prediction from lidar latent codes
phenos_list <- c("silking_days", "anthesis_days", "grain_yield", "plant_height", "ear_height", "stand_count", "stalk_lodging", "grain_moisture", "ASI")
pred_results_lidar <- c()
for(pheno in phenos_list){
  set.seed(10)
  print(pheno)
  #get phenotype
  phenoDat <- df_lidar[,c("uniqueID", pheno, "genotype_num")]
  colnames(phenoDat) <- c("uniqueID", "pheno", "genotype_num")
  #remove NAs from phenoDat
  phenoDat <- phenoDat[!is.na(phenoDat$pheno),]
  #get phenotype for each uniqueID
  phenoDat <- phenoDat %>% group_by_at(vars(uniqueID)) %>%
    summarise(across(where(is.numeric), ~mean(.x, na.rm = TRUE)), across(!where(is.numeric), ~first(.x)))
  #remove uniqueIDs where pheno is NA from cov mat
  cov_mat <- lidar_timeseries_scaled_by_lf[(rownames(lidar_timeseries_scaled_by_lf) %in% phenoDat$uniqueID), ]
  #sort pheno by uniqueID of cov mat & remove uniqueIDs not in cov mat
  rownames(phenoDat) <- phenoDat$uniqueID
  phenoDat <- phenoDat[rownames(cov_mat), ]
  #mean center phenotype by year
  sub_2020 <- grepl("2020", phenoDat$uniqueID)
  sub_2021 <- grepl("2021", phenoDat$uniqueID)
  sub_2022 <- grepl("2022", phenoDat$uniqueID)
  sub_2023 <- grepl("2023", phenoDat$uniqueID)
  mean_val_2020 <- mean(phenoDat$pheno[sub_2020], na.rm = TRUE)
  mean_val_2021 <- mean(phenoDat$pheno[sub_2021], na.rm = TRUE)
  mean_val_2022 <- mean(phenoDat$pheno[sub_2022], na.rm = TRUE)
  mean_val_2023 <- mean(phenoDat$pheno[sub_2023], na.rm = TRUE)
  phenoDat$pheno[sub_2020] <- phenoDat$pheno[sub_2020] - mean_val_2020
  phenoDat$pheno[sub_2021] <- phenoDat$pheno[sub_2021] - mean_val_2021
  phenoDat$pheno[sub_2022] <- phenoDat$pheno[sub_2022] - mean_val_2022
  phenoDat$pheno[sub_2023] <- phenoDat$pheno[sub_2023] - mean_val_2023
  for(i in 1:10){
    save_at <- paste0("BGLR_output/CV1_cv_comparisons/lidar_", pheno, "_", cv, "_", lsp_size, "_", i, "_")
    bglr_mod <- run_BGLR(cov_mat, phenoDat, save_at)
    pred_results <- data.frame(CV = cv, LatentDimSize = lsp_size, EquivDates = equivdates, Pheno = pheno, Predictor = "LiDAR Latent Phenotypes", Correlation = bglr_mod[1], RMSE = bglr_mod[2])
    # pred_results <- bglr_mod
    # pred_results$pheno <- pheno
    pred_results_lidar <- rbind(pred_results_lidar, pred_results)
  }
}

write.csv(pred_results_lidar, paste0("pred_results/CV_comparisons/pred_lidar_", cv, "_", lsp_size, ".csv"))

#Prediction from integrated latent codes
phenos_list <- c("silking_days", "anthesis_days", "grain_yield", "plant_height", "ear_height", "stand_count", "stalk_lodging", "grain_moisture", "ASI")
pred_results_int <- c()
for(pheno in phenos_list){
  set.seed(10)
  print(pheno)
  #get phenotype
  phenoDat <- df_msi[,c("uniqueID", pheno, "genotype_num")]
  colnames(phenoDat) <- c("uniqueID", "pheno", "genotype_num")
  #remove NAs from phenoDat
  phenoDat <- phenoDat[!is.na(phenoDat$pheno),]
  #get phenotype for each uniqueID
  phenoDat <- phenoDat %>% group_by_at(vars(uniqueID)) %>%
    summarise(across(where(is.numeric), ~mean(.x, na.rm = TRUE)), across(!where(is.numeric), ~first(.x)))
  #remove uniqueIDs where pheno is NA from cov mat
  cov_mat <- integrated_timeseries_scaled_by_lf[(rownames(integrated_timeseries_scaled_by_lf) %in% phenoDat$uniqueID), ]
  #sort pheno by uniqueID of cov mat & remove uniqueIDs not in cov mat
  rownames(phenoDat) <- phenoDat$uniqueID
  phenoDat <- phenoDat[rownames(cov_mat), ]
  #mean center phenotype by year
  sub_2020 <- grepl("2020", phenoDat$uniqueID)
  sub_2021 <- grepl("2021", phenoDat$uniqueID)
  sub_2022 <- grepl("2022", phenoDat$uniqueID)
  sub_2023 <- grepl("2023", phenoDat$uniqueID)
  mean_val_2020 <- mean(phenoDat$pheno[sub_2020], na.rm = TRUE)
  mean_val_2021 <- mean(phenoDat$pheno[sub_2021], na.rm = TRUE)
  mean_val_2022 <- mean(phenoDat$pheno[sub_2022], na.rm = TRUE)
  mean_val_2023 <- mean(phenoDat$pheno[sub_2023], na.rm = TRUE)
  phenoDat$pheno[sub_2020] <- phenoDat$pheno[sub_2020] - mean_val_2020
  phenoDat$pheno[sub_2021] <- phenoDat$pheno[sub_2021] - mean_val_2021
  phenoDat$pheno[sub_2022] <- phenoDat$pheno[sub_2022] - mean_val_2022
  phenoDat$pheno[sub_2023] <- phenoDat$pheno[sub_2023] - mean_val_2023
  for(i in 1:10){
    save_at <- paste0("BGLR_output/CV1_dtype_comparisons/int_", pheno, "_", cv, "_", lsp_size, "_", i, "_")
    bglr_mod <- run_BGLR(cov_mat, phenoDat, save_at)
    pred_results <- data.frame(CV = cv, LatentDimSize = lsp_size, EquivDates = equivdates, Pheno = pheno, Predictor = "Integrated Latent Phenotypes", Correlation = bglr_mod[1], RMSE = bglr_mod[2])
    # pred_results <- bglr_mod
    # pred_results$pheno <- pheno
    pred_results_int <- rbind(pred_results_int, pred_results)
  }
}

write.csv(pred_results_int, paste0("pred_results/CV_dtype_comparisons/pred_int_", cv, "_", lsp_size, ".csv"))

pred_results_all <- rbind(pred_results_msi, pred_results_lidar, pred_results_int)

write.csv(pred_results_all, "pred_results/pred_all_", cv, "_", lsp_size, ".csv")
```


#Model Comparison - Predict Pheno from LSPs using Bayesian Regression - Season-Level Single train/val Split
```{r}
run_BGLR_1fold <- function(cov_matrix, phenoDat, saveAt){
  #get val uniqueIDs
  val <- unique(phenoDat$uniqueID[phenoDat$train_set == 'val'])
  Y <- phenoDat
  #mask data for prediction
  Y$pheno[Y$uniqueID %in% val] <- NA
  Y = as.matrix(Y$pheno)
  #set up ETA list
  Model <- list(list(X = cov_matrix, model = "BRR"))
  result <- BGLR(Y, response_type = "gaussian", ETA = Model, nIter = 40000, burnIn = 500, thin = 200, verbose = FALSE, saveAt = saveAt)
  #get pheno estimates from result list
  Yhat = result$yHat
  #get df of pred and actual phenotypes
  df = data.frame(pred = Yhat, actual = as.matrix(phenoDat$pheno), uniqueID = as.matrix(phenoDat$uniqueID))
  #get train uniqueIDs
  train <- unique(phenoDat$uniqueID[phenoDat$train_set == 'train'])
  #correlation between training set pred and actual
  cor_train = cor(df$pred[df$uniqueID %in% train], df$actual[df$uniqueID %in% train])
  print(paste0("training set accuracy is: ", cor_train))
  #correlation between validation set pred and actual
  cor = cor(df$pred[df$uniqueID %in% val], df$actual[df$uniqueID %in% val])
  print(paste0("validation set accuracy is: ", cor))
  #rmse between validation set pred and actual
  rmse = sqrt(mean((df$actual[df$uniqueID %in% val] - df$pred[df$uniqueID %in% val])^2))
  print(rmse)
  #return correlation and rmse
  return(c(cor, rmse))
  #return(df)
}

#Prediction from VIs
phenos_list <- c("silking_days", "anthesis_days", "grain_yield", "plant_height", "ear_height", "stand_count", "stalk_lodging", "grain_moisture", "ASI")
pred_results_vi <- c()
for(pheno in phenos_list){
  set.seed(10)
  print(pheno)
  #get phenotype
  phenoDat <- df_msi[,c("uniqueID", pheno, "train_set")]
  colnames(phenoDat) <- c("uniqueID", "pheno", "train_set")
  #remove NAs from phenoDat
  phenoDat <- phenoDat[!is.na(phenoDat$pheno),]
  #get phenotype for each uniqueID
  phenoDat <- phenoDat %>%
    distinct(uniqueID, .keep_all = TRUE)
  #remove uniqueIDs where pheno is NA from cov mat
  cov_mat <- vi_timeseries_scaled_by_lf[(rownames(vi_timeseries_scaled_by_lf) %in% phenoDat$uniqueID), ]
  #sort pheno by uniqueID of cov mat & remove uniqueIDs not in cov mat
  rownames(phenoDat) <- phenoDat$uniqueID
  phenoDat <- phenoDat[rownames(cov_mat), ]
  #mean center phenotype by year
  sub_2020 <- grepl("2020", phenoDat$uniqueID)
  sub_2021 <- grepl("2021", phenoDat$uniqueID)
  sub_2022 <- grepl("2022", phenoDat$uniqueID)
  sub_2023 <- grepl("2023", phenoDat$uniqueID)
  sub_2024_NYH2 <- grepl("NYH2_2024", phenoDat$uniqueID)
  sub_2024_NYH3 <- grepl("NYH3_2024", phenoDat$uniqueID)
  mean_val_2020 <- mean(phenoDat$pheno[sub_2020], na.rm = TRUE)
  mean_val_2021 <- mean(phenoDat$pheno[sub_2021], na.rm = TRUE)
  mean_val_2022 <- mean(phenoDat$pheno[sub_2022], na.rm = TRUE)
  mean_val_2023 <- mean(phenoDat$pheno[sub_2023], na.rm = TRUE)
  mean_val_2024_NYH2 <- mean(phenoDat$pheno[sub_2024_NYH2], na.rm = TRUE)
  mean_val_2024_NYH3 <- mean(phenoDat$pheno[sub_2024_NYH3], na.rm = TRUE)
  phenoDat$pheno[sub_2020] <- phenoDat$pheno[sub_2020] - mean_val_2020
  phenoDat$pheno[sub_2021] <- phenoDat$pheno[sub_2021] - mean_val_2021
  phenoDat$pheno[sub_2022] <- phenoDat$pheno[sub_2022] - mean_val_2022
  phenoDat$pheno[sub_2023] <- phenoDat$pheno[sub_2023] - mean_val_2023
  phenoDat$pheno[sub_2024_NYH2] <- phenoDat$pheno[sub_2024_NYH2] - mean_val_2024_NYH2
  phenoDat$pheno[sub_2024_NYH3] <- phenoDat$pheno[sub_2024_NYH3] - mean_val_2024_NYH3
  save_at <- paste0("BGLR_output/Test_dtype_comparisons/VI_", pheno, "_", cv, cv_version, "_")
  bglr_mod <- run_BGLR_1fold(cov_mat, phenoDat, save_at)
  pred_results <- data.frame(CV = cv, Version = cv_version, LatentDimSize = lsp_size, EquivDates = equivdates, Pheno = pheno, Predictor = "Vegetation Indices", Correlation = bglr_mod[1], RMSE = bglr_mod[2])
  # pred_results <- bglr_mod
  # pred_results$pheno <- pheno
  pred_results_vi <- rbind(pred_results_vi, pred_results)
}

write.csv(pred_results_vi, paste0("pred_results/CV_dtype_comparisons/pred_VI_", cv, cv_version, ".csv"))

#Prediction from MSI latent codes
phenos_list <- c("silking_days", "anthesis_days", "grain_yield", "plant_height", "ear_height", "stand_count", "stalk_lodging", "grain_moisture", "ASI")
pred_results_msi <- c()
for(pheno in phenos_list){
  set.seed(10)
  print(pheno)
  #get phenotype
  phenoDat <- df_msi[,c("uniqueID", pheno, "train_set")]
  colnames(phenoDat) <- c("uniqueID", "pheno", "train_set")
  #remove NAs from phenoDat
  phenoDat <- phenoDat[!is.na(phenoDat$pheno),]
  #get phenotype for each uniqueID
  phenoDat <- phenoDat %>%
    distinct(uniqueID, .keep_all = TRUE)
  #remove uniqueIDs where pheno is NA from cov mat
  cov_mat <- msi_timeseries_scaled_by_lf[(rownames(msi_timeseries_scaled_by_lf) %in% phenoDat$uniqueID), ]
  #sort pheno by uniqueID of cov mat & remove uniqueIDs not in cov mat
  rownames(phenoDat) <- phenoDat$uniqueID
  phenoDat <- phenoDat[rownames(cov_mat), ]
  #mean center phenotype by year
  sub_2020 <- grepl("2020", phenoDat$uniqueID)
  sub_2021 <- grepl("2021", phenoDat$uniqueID)
  sub_2022 <- grepl("2022", phenoDat$uniqueID)
  sub_2023 <- grepl("2023", phenoDat$uniqueID)
  sub_2024_NYH2 <- grepl("NYH2_2024", phenoDat$uniqueID)
  sub_2024_NYH3 <- grepl("NYH3_2024", phenoDat$uniqueID)
  mean_val_2020 <- mean(phenoDat$pheno[sub_2020], na.rm = TRUE)
  mean_val_2021 <- mean(phenoDat$pheno[sub_2021], na.rm = TRUE)
  mean_val_2022 <- mean(phenoDat$pheno[sub_2022], na.rm = TRUE)
  mean_val_2023 <- mean(phenoDat$pheno[sub_2023], na.rm = TRUE)
  mean_val_2024_NYH2 <- mean(phenoDat$pheno[sub_2024_NYH2], na.rm = TRUE)
  mean_val_2024_NYH3 <- mean(phenoDat$pheno[sub_2024_NYH3], na.rm = TRUE)
  phenoDat$pheno[sub_2020] <- phenoDat$pheno[sub_2020] - mean_val_2020
  phenoDat$pheno[sub_2021] <- phenoDat$pheno[sub_2021] - mean_val_2021
  phenoDat$pheno[sub_2022] <- phenoDat$pheno[sub_2022] - mean_val_2022
  phenoDat$pheno[sub_2023] <- phenoDat$pheno[sub_2023] - mean_val_2023
  phenoDat$pheno[sub_2024_NYH2] <- phenoDat$pheno[sub_2024_NYH2] - mean_val_2024_NYH2
  phenoDat$pheno[sub_2024_NYH3] <- phenoDat$pheno[sub_2024_NYH3] - mean_val_2024_NYH3
  save_at <- paste0("BGLR_output/Test_dtype_comparisons/msi_", pheno, "_", cv, cv_version, "_", lsp_size, "_")
  bglr_mod <- run_BGLR_1fold(cov_mat, phenoDat, save_at)
  pred_results <- data.frame(CV = cv, Version = cv_version, LatentDimSize = lsp_size, EquivDates = equivdates, Pheno = pheno, Predictor = "MSI Latent Phenotypes", Correlation = bglr_mod[1], RMSE = bglr_mod[2])
  # pred_results <- bglr_mod
  # pred_results$pheno <- pheno
  pred_results_msi <- rbind(pred_results_msi, pred_results)
}

write.csv(pred_results_msi, paste0("pred_results/CV_dtype_comparisons/pred_msi_", cv, cv_version, "_", lsp_size, ".csv"))

#Prediction from lidar latent codes
phenos_list <- c("silking_days", "anthesis_days", "grain_yield", "plant_height", "ear_height", "stand_count", "stalk_lodging", "grain_moisture", "ASI")
pred_results_lidar <- c()
for(pheno in phenos_list){
  set.seed(10)
  print(pheno)
  #get phenotype
  phenoDat <- df_lidar[,c("uniqueID", pheno, "train_set")]
  colnames(phenoDat) <- c("uniqueID", "pheno", "train_set")
  #remove NAs from phenoDat
  phenoDat <- phenoDat[!is.na(phenoDat$pheno),]
  #get phenotype for each uniqueID
  phenoDat <- phenoDat %>%
    distinct(uniqueID, .keep_all = TRUE)
  #remove uniqueIDs where pheno is NA from cov mat
  cov_mat <- lidar_timeseries_scaled_by_lf[(rownames(lidar_timeseries_scaled_by_lf) %in% phenoDat$uniqueID), ]
  #sort pheno by uniqueID of cov mat & remove uniqueIDs not in cov mat
  rownames(phenoDat) <- phenoDat$uniqueID
  phenoDat <- phenoDat[rownames(cov_mat), ]
  #mean center phenotype by year
  sub_2020 <- grepl("2020", phenoDat$uniqueID)
  sub_2021 <- grepl("2021", phenoDat$uniqueID)
  sub_2022 <- grepl("2022", phenoDat$uniqueID)
  sub_2023 <- grepl("2023", phenoDat$uniqueID)
  sub_2024_NYH2 <- grepl("NYH2_2024", phenoDat$uniqueID)
  sub_2024_NYH3 <- grepl("NYH3_2024", phenoDat$uniqueID)
  mean_val_2020 <- mean(phenoDat$pheno[sub_2020], na.rm = TRUE)
  mean_val_2021 <- mean(phenoDat$pheno[sub_2021], na.rm = TRUE)
  mean_val_2022 <- mean(phenoDat$pheno[sub_2022], na.rm = TRUE)
  mean_val_2023 <- mean(phenoDat$pheno[sub_2023], na.rm = TRUE)
  mean_val_2024_NYH2 <- mean(phenoDat$pheno[sub_2024_NYH2], na.rm = TRUE)
  mean_val_2024_NYH3 <- mean(phenoDat$pheno[sub_2024_NYH3], na.rm = TRUE)
  phenoDat$pheno[sub_2020] <- phenoDat$pheno[sub_2020] - mean_val_2020
  phenoDat$pheno[sub_2021] <- phenoDat$pheno[sub_2021] - mean_val_2021
  phenoDat$pheno[sub_2022] <- phenoDat$pheno[sub_2022] - mean_val_2022
  phenoDat$pheno[sub_2023] <- phenoDat$pheno[sub_2023] - mean_val_2023
  phenoDat$pheno[sub_2024_NYH2] <- phenoDat$pheno[sub_2024_NYH2] - mean_val_2024_NYH2
  phenoDat$pheno[sub_2024_NYH3] <- phenoDat$pheno[sub_2024_NYH3] - mean_val_2024_NYH3
  save_at <- paste0("BGLR_output/Test_dtype_comparisons/lidar_", pheno, "_", cv, cv_version, "_", lsp_size, "_")
  bglr_mod <- run_BGLR_1fold(cov_mat, phenoDat, save_at)
  pred_results <- data.frame(CV = cv, Version = cv_version, LatentDimSize = lsp_size, EquivDates = equivdates, Pheno = pheno, Predictor = "LiDAR Latent Phenotypes", Correlation = bglr_mod[1], RMSE = bglr_mod[2])
  # pred_results <- bglr_mod
  # pred_results$pheno <- pheno
  pred_results_lidar <- rbind(pred_results_lidar, pred_results)
}

write.csv(pred_results_lidar, paste0("pred_results/CV_dtype_comparisons/pred_lidar_", cv, cv_version, "_", lsp_size, ".csv"))

#Prediction from integrated latent codes
phenos_list <- c("silking_days", "anthesis_days", "grain_yield", "plant_height", "ear_height", "stand_count", "stalk_lodging", "grain_moisture", "ASI")
pred_results_int <- c()
for(pheno in phenos_list){
  set.seed(10)
  print(pheno)
  #get phenotype
  phenoDat <- df_msi[,c("uniqueID", pheno, "train_set")]
  colnames(phenoDat) <- c("uniqueID", "pheno", "train_set")
  #remove NAs from phenoDat
  phenoDat <- phenoDat[!is.na(phenoDat$pheno),]
  #get phenotype for each uniqueID
  phenoDat <- phenoDat %>%
    distinct(uniqueID, .keep_all = TRUE)
  #remove uniqueIDs where pheno is NA from cov mat
  cov_mat <- integrated_timeseries_scaled_by_lf[(rownames(integrated_timeseries_scaled_by_lf) %in% phenoDat$uniqueID), ]
  #sort pheno by uniqueID of cov mat & remove uniqueIDs not in cov mat
  rownames(phenoDat) <- phenoDat$uniqueID
  phenoDat <- phenoDat[rownames(cov_mat), ]
  #mean center phenotype by year
  sub_2020 <- grepl("2020", phenoDat$uniqueID)
  sub_2021 <- grepl("2021", phenoDat$uniqueID)
  sub_2022 <- grepl("2022", phenoDat$uniqueID)
  sub_2023 <- grepl("2023", phenoDat$uniqueID)
  sub_2024_NYH2 <- grepl("NYH2_2024", phenoDat$uniqueID)
  sub_2024_NYH3 <- grepl("NYH3_2024", phenoDat$uniqueID)
  mean_val_2020 <- mean(phenoDat$pheno[sub_2020], na.rm = TRUE)
  mean_val_2021 <- mean(phenoDat$pheno[sub_2021], na.rm = TRUE)
  mean_val_2022 <- mean(phenoDat$pheno[sub_2022], na.rm = TRUE)
  mean_val_2023 <- mean(phenoDat$pheno[sub_2023], na.rm = TRUE)
  mean_val_2024_NYH2 <- mean(phenoDat$pheno[sub_2024_NYH2], na.rm = TRUE)
  mean_val_2024_NYH3 <- mean(phenoDat$pheno[sub_2024_NYH3], na.rm = TRUE)
  phenoDat$pheno[sub_2020] <- phenoDat$pheno[sub_2020] - mean_val_2020
  phenoDat$pheno[sub_2021] <- phenoDat$pheno[sub_2021] - mean_val_2021
  phenoDat$pheno[sub_2022] <- phenoDat$pheno[sub_2022] - mean_val_2022
  phenoDat$pheno[sub_2023] <- phenoDat$pheno[sub_2023] - mean_val_2023
  phenoDat$pheno[sub_2024_NYH2] <- phenoDat$pheno[sub_2024_NYH2] - mean_val_2024_NYH2
  phenoDat$pheno[sub_2024_NYH3] <- phenoDat$pheno[sub_2024_NYH3] - mean_val_2024_NYH3
  save_at <- paste0("BGLR_output/Test_dtype_comparisons/int_", pheno, "_", cv, cv_version, "_", lsp_size, "_")
  bglr_mod <- run_BGLR_1fold(cov_mat, phenoDat, save_at)
  pred_results <- data.frame(CV = cv, Version = cv_version, LatentDimSize = lsp_size, EquivDates = equivdates, Pheno = pheno, Predictor = "Integrated Latent Phenotypes", Correlation = bglr_mod[1], RMSE = bglr_mod[2])
  # pred_results <- bglr_mod
  # pred_results$pheno <- pheno
  pred_results_int <- rbind(pred_results_int, pred_results)
}

write.csv(pred_results_int, paste0("pred_results/CV_dtype_comparisons/pred_int_", cv, cv_version, "_", lsp_size, ".csv"))
```


